### 2014-05-29 ACL2014/notes.tex
Added some notes in system.tex

Most of the work done in LSH in NLP community uses LSH for dimensionality reduction.

\section{Contributions}
\begin{itemize}
\item We present a distributed system that can approximately compute pairwise similarity 
between all queries on a commercial large-scale query logs in a time efficient manner. 
Can we claim something more specific wrt time (e.g. 0(N logN))?
\item Vanilla LSH 
\item We propose two novel variants of vanilla LSH motivated by the research on Multi-Probe LSH \cite{LvVLDB07}: Random Flipping and Query-Driven Flipping. 
\item  We show that the applicability of our system on two real-world applications like finding related queries (Increasing  direct display coverage is related to finding related queries) and ??.
\end{itemize}

\section{Observations}
\begin{itemize}
\item How to vary $K$ and $L$? The selection of $K$ and $L$ is data dependent. 
If we increase $K$ (increasing the precision to reduce false positives), we also need to 
increase $L$ to get good recall (increasing the recall to reduce false negatives). 
These parameters can be tuned for a specific dataset/application by randomly sampling small set of queries.\footnote{In this paper, we randomly selected 2000 queries to tune $K$ and $L$ parameters.}
\item

\end{itemize}

Advantages of Flipping bits compared to increasing L:
1. Increasing $L$ means generating more random projection bits that is both time and memory intensive.   
%Note, we are already using the trick mentioned in Original LSH to decrease
%the number of tables. Our implementation requires O(sqrt(L)) tables.
2. Probing in more tables for vanilla LSH, hence we need to read data from the disk (more disk read operations.). While Flipping bits, data is
already in memory, and the given data needs to be flipped.
3. Disadvantage of flipping is we need to store random projection distances to perform flipping (which takes space).