### 2014-05-30 ACL2014/abstract.tex
Updated latex files

\begin{abstract}
Many Web, Mobile, and NLP applications 
rely on finding nearest neighbors 
for a given query from a commercial large-scale query logs . 
Due to the large scale volume of queries, 
computing nearest neighbors from all queries using brute force linear 
search is a computationally time intensive task. 
To challenge the aforementioned problem, we exploit existing 
\lshf (a.k.a, \lsh) methods and propose 
their novel variants in distributed setting (e.g. Hadoop). 
Our experimental results show that couple of our 
new variants of \lsh get significantly better recall than vanilla \lsh by 
using the same number of hash tables. 
\end{abstract}