### 2014-05-31 ACL2014/applications.tex

\section{Applications}
We show the applicability of our nearest neighbors geneated using the approximate \dbflip algorithm with parameters $L=10$, $K=24$, and $F=2$ on two well-known query understanding tasks.