### 2014-05-31 ACL2014/applications.tex
Added updated latex file

\section{Applications}
We show the applicability of our nearest neighbors geneated using 
the approximate \dflip algorithm with parameters $L=10$, $K=24$, and $F=2$ on 
two well-known query understanding tasks: finding related queries and removing de-duplcate queries.