### 2014-06-01 ACL2014/discussion.tex
updated latex files

\section{Discussion and Conclusion}
\label{sec:discuss}
%In this paper, we demonstrated the effectiveness of our p