### 2014-06-03 ACL2014/relatedwork.tex
related work

\section{Related Work}
Locality sensitive hashing is a very well studied algorithmic tool, that has found practical uses too. 
In processing large text corpora, however, LSH algorithms have been less used, primarily because of their
still significant memory requirements. Recent work by~\cite{zadeh2013dimension} developed a modification of LSH algorithm, for 
the Jaccard distance metric, that scales to very large text corpora by virtue of being implemented on a map-reduce, 
and by using clever sampling schemes in order to reduce the communication cost. Our work is for the cosine similarity metric, 
and uses bit flipping in a distributed manner to reduce the number of hash tables in LSH and hence the memory. 
Other innovations to LSH methods in the text corpora analysis domain include developing methods to compute
LSH signatures in an online manner~\cite{van2010online}, and using reservoir sampling to compute the signatures~\cite{van2011efficient}.